\chapter{System Design and Implementation}
This chapter provides a detailed technical overview of the system architecture and the specific software implementations developed to support parallel optical links. It outlines the extensive modifications made to the TeraFlowSDN ecosystem, the creation of a standalone RSA computation module, and the deployment of the novel ParallelOpticalController microservice.
\section{Overall System Architecture}
The proposed solution is built entirely within the TeraFlowSDN (TFS) microservices ecosystem. TFS operates as a cloud-native, Kubernetes-based SDN controller, making it highly modular and extensible. To solve the Routing and Spectrum Allocation (RSA) problem for parallel links, the system architecture was augmented with a newly developed ParallelOpticalController service. This service interfaces heavily with the existing ContextService (for retrieving topology and endpoint database states) and the DeviceService (for communicating with the physical and emulated optical hardware). The architecture ensures that lightpath requests are intercepted by the ParallelOpticalController, which computes the optimal routing and spectrum allocation before instructing the DeviceService to push XML configurations to the underlying transponders and ROADMs.
\section{Modifications to TeraFlowSDN}
\subsection{Device Service Enhancements (OCDriver, OpenROADMDriver)}
To perform accurate device-based spectrum filtration, the controller requires detailed hardware telemetry. The existing OCDriver (for OpenConfig devices) was heavily modified to extract device information directly from the device's XML configuration files rather than relying solely on static JSON descriptors.
\begin{enumerate}
    \item A new device-specific filter mechanism was introduced, establishing a  device\_filters package containing classes like DefaultDevice.py and CNITTestROADMDevice.py.
    \item The OCDriver.py was updated to execute a two-phase extraction process: it first retrieves the model\_name from the device, and then maps this model to the appropriate filter class to extract transponder and port values accurately.
    \item Similar enhancements were made to the OpenROADMDriver, filtering out internal mapping ports and correctly marking client ports as OCH transport types and line ports as OMS transport types.
\end{enumerate}
\subsection{Context Service \& Database Schema Extensions}
Because the original database lacked fields for manufacturer and hardware specifications, significant schema extensions were introduced to ContextService.
\begin{enumerate}
    \item The DeviceModel.py table was updated to include mandatory fields: mfg\_name, model, and serial\_no, alongside nullable fields for hardware\_version and software\_version.
    \item The EndPointModel.py was extended to include an index string and a transport\_type. The index acts as a unique identifier (e.g., "TP1\_11") generated by combining a sanitized device name and the port index, resolving previous endpoint mapping ambiguities.
    \item The OpticalChannelModel was upgraded to store crucial RSA metrics: \texttt{min\_frequency}, \texttt{max\_frequency}, \texttt{flex\_slots}, and \texttt{bitmap\_value}. These schema changes were formally defined in the context.proto protocol buffers and propagated throughout the gRPC services.
\end{enumerate}
\section{RSA Computation Module (RSATools, ITUStandards)}
To centralize the mathematical and logical rules of optical networking, two primary Python helper modules were developed: ITUStandards.py and RSATools.py.
\begin{itemize}
    \item \textbf{ITUStandards.py:} This module codifies the strict ITU-T G.694.1 standards. It defines constants such as the 193.1 THz anchor frequency and the 6.25 GHz slot granularity. It also encapsulates enumerations for frequency ranges across all optical bands (O, E, S, C, L) and slot availability statuses (AVAILABLE, UNAVAILABLE, \texttt{IN\_USE}).
    \item \textbf{RSATools.py:} Functioning as the core logic engine, this module handles the dynamic RSA calculations. It includes functions like detect\_band() to identify the smallest optical band containing a given frequency, and compute\_rsa\_params() to determine the minimum and maximum frequencies and required flex slots. Crucially, it calculates the base hardware spectrum bitmap using the formula $B_{\text{map}} = 2^{\text{slots}} - 1$ and dynamically generates spectrum slot tables for both routing computations and WebUI visualization.
\end{itemize}
\section{Parallel Optical Controller Microservice}
\subsection{Service Architecture \& API Design}
The ParallelOpticalController was engineered as an autonomous microservice within TFS. It is containerized using Docker and deployed into the Kubernetes cluster alongside existing services. Following the deployment, both the serviceservice and webuiservice undergo rollout restarts to discover the new controller's environment variables and navigation endpoints.
\subsection{Graph Construction with NetworkX}
When an optical intent is received, the controller does not rely on static routing tables; instead, it dynamically reconstructs the network graph. Fetching endpoint and optical link data via the Context client, it utilizes the networkX package to build complex multi-directed graphs. The module \texttt{topology.py} is responsible for defining graph nodes (active optical devices) and generating distinct graph models, such as \texttt{G\_FREE} for links where OCH is unused, and \texttt{G\_simple} for raw path discovery.
\section{Integration with Existing TFS Services}
The controller interfaces seamlessly with the ContextService to read the current network state. During the "Slot Acquisition" phase, the controller updates the flow parameters in the database via the Context client, ensuring the allocated bitwise spectrum blocks are officially locked and recorded.
\section{Topology \& Link Management (Parallel Link Support)}
A major implementation hurdle was supporting topologies with parallel links natively. The traditional method of mapping links via \texttt{endpoint\_uuid} often failed when deploying topologies dynamically due to hashing mismatches. To resolve this, link descriptors were upgraded to support an \texttt{endpoint\_index} fallback mechanism. Furthermore, to maintain a clean graphical user interface without overlapping link lines, parallel links are grouped and visually represented using an integer multiplier format, such as xN, detailing the exact number of parallel connections existing between two adjacent devices.
\section{WebUI Enhancements (Spectrum Visualization)}
To expose the advanced capabilities of the backend, the TFS Web User Interface underwent significant improvements.
\begin{itemize}
    \item The device details template (detail.html) was modified to include the newly extracted Index and Transport Type metrics.
    \item A dedicated "Endpoint Frequency View" template was engineered, featuring a comprehensive "Frequency Capabilities Card" that displays wavelength ranges, bandwidth, and the specific optical band name.
    \item Most notably, the WebUI now features a real-time, scrollable "Spectrum Allocation Table". This table displays every individual 6.25 GHz slot within the device's band, utilizing color-coded badges to indicate slot status: green for AVAILABLE, yellow for \texttt{IN\_USE}, gray for UNAVAILABLE (outside hardware bounds), and blue for the standard ANCHOR frequency.
\end{itemize}