\chapter{Background and Literature Review}
This chapter establishes the theoretical foundation and contextual background for the research presented in this thesis. It reviews the evolution of Software-Defined Networking (SDN) in optical domains, details the architectural components of TeraFlowSDN and Optical Transport Networks (OTN), and provides a critical analysis of current Routing and Spectrum Allocation (RSA) methodologies. Finally, it identifies the limitations of existing approaches when handling parallel optical links, thereby setting the stage for the proposed solution.
\section{Software Defined Network}
\subsection{SDN Architecture and Principles}
Software-Defined Networking (SDN) represents a paradigm shift in telecommunications, decoupling the network's control logic (the control plane) from the underlying packet or optical forwarding hardware (the data plane). This separation centralizes network intelligence into a software-based controller, which maintains a global view of the network topology. The core principles of SDN include centralized management, network programmability, and the use of open, standardized APIs (such as OpenFlow or NETCONF) to dictate forwarding behaviors dynamically.
\subsection{SDN in Optical Networks}
While initially designed for packet-switched networks, SDN principles have been aggressively adapted for optical networks to create Software-Defined Optical Networks (SDON). Unlike packet networks, optical networks deal with analog physical layer impairments, wavelength continuity, and spectrum allocation. SDON abstracts these physical layer complexities, allowing the central controller to dynamically provision lightpaths, tune transceivers, and configure Reconfigurable Optical Add-Drop Multiplexers (ROADMs) automatically, moving away from historically static, manually operated optical grids.
\section{TeraflowSDN}
\subsection{Architecture and Ecosystem}
TeraFlowSDN (TFS) is a modern, cloud-native SDN controller designed to manage complex, multi-tenant optical and packet networks. Built upon a microservices architecture, TFS operates within Kubernetes environments, allowing individual control components to scale independently. This modularity makes it an ideal ecosystem for integrating custom logic, such as vendor-agnostic device drivers and novel path computation engines.
\subsection{Key services}
The TFS ecosystem relies on several critical microservices to maintain network state and execute configurations. The Context Service manages the database schemas, storing topologies, devices, endpoints, and optical link data. The Device Service handles south-bound communications, utilizing specific drivers to interact with physical or emulated hardware. Furthermore, the OpticalController acts as a dedicated service to handle optical intents, supported by a graphical WebUI for visual management and telemetry display.
\section{Optical Transport Networks}
\subsection{WDM technology}
Wavelength Division Multiplexing (WDM), specifically Dense WDM (DWDM), is the foundational technology for modern high-capacity optical networks. DWDM enables the transmission of multiple data streams over a single optical fiber by modulating each stream onto a distinct wavelength (or color) of laser light . Historically, DWDM utilized a fixed-grid spacing (e.g., 50 GHz or 100 GHz); however, modern elastic optical networks (EONs) utilize flexible-grid (flex-grid) technologies to allocate variable bandwidth depending on the exact requirements of the traffic.
\subsection{ITU standards}
The implementation of flex-grid networks is strictly governed by International Telecommunication Union (ITU) standards. Specifically, the ITU-T G.694.1 standard defines the spectral grid for DWDM applications. Under this standard, the spectrum is anchored at a central frequency of 193.1 THz. The available bandwidth is divided into minimal, indivisible frequency slots with a granularity of 6.25 GHz. Spectrum allocation must adhere to this granularity, operating across defined optical bands (such as the C, L, and S bands).
\subsection{Network elements (devices)}
The physical topology comprises two primary types of nodes: transponders (and muxponders) and ROADMs. Transponders map client-layer signals into optical channels suitable for DWDM transmission, handling modulation and forward error correction (FEC). ROADMs, conversely, are optical switches that allow individual wavelengths to be added, dropped, or passed through a node dynamically without requiring optical-to-electrical-to-optical (O-E-O) conversion.
\subsection{Data models (configuration)}
To achieve vendor interoperability, SDON relies on standardized YANG data models. OpenConfig and OpenROADM are two prominent models used to standardize the representation of optical device configurations and operational states. These models define how parameters—such as target output power, operational frequencies, and transport types (e.g., OCH, OMS)—are structured in XML or JSON formats, allowing the SDN controller to parse and push configurations uniformly regardless of the equipment manufacturer.
\section{Routing and Spectrum Allocation}
\subsection{Spectrum continuity and contiguity constraints}
Establishing a lightpath in an EON requires solving the Routing and Spectrum Allocation (RSA) problem. RSA is constrained by two strict physical requirements:

\begin{enumerate}
    \item Spectrum Continuity: The exact same optical frequency must be maintained across all hops of a multi-hop route, provided no wavelength converters are present.
    \item Spectrum Contiguity: The frequency slots assigned to a single lightpath must be adjacent to one another within the optical spectrum to form a contiguous bandwidth block.
\end{enumerate}
\subsection{Existing RSA algorithms}
TBD
\section{Problems with Parallel Links in RSA}
\subsection{Problem statement}
While standard RSA algorithms perform adequately on simple topologies, their computational efficiency degrades drastically in multi-graph networks featuring multiple parallel optical links between the same adjacent devices. In such topologies, the total number of possible path combinations from a source to a destination increases exponentially. Because signals routed over the same degree of a ROADM cannot reuse identical frequencies without severe interference, the controller must exhaustively calculate and verify the availability of every 6.25 GHz slot across every parallel edge. This results in massive computational overhead and latency during lightpath provisioning.
\section{Current approaches and limitation}
Contemporary SDN controllers, such as ONOS and OpenDaylight, lack native optimizations for multi-directed parallel optical graphs. Currently, these systems treat parallel edges as entirely isolated, individual channels. When computing a path, the RSA algorithm evaluates each parallel link sequentially, executing redundant continuity and contiguity checks. This back-and-forth, iterative calculation methodology is inherently unscalable and fails to leverage the collective capacity of parallel optical pipes efficiently.
\section{Related work}
A substantial body of literature exists on optimizing RSA in Elastic Optical Networks (EONs). Previous research has extensively modeled multi-hop routing constraints, fragmentation minimization, and modulation format adaptation. However, literature addressing the specific algorithmic handling of parallel links within a centralized, vendor-agnostic SDN architecture remains sparse. Most existing works assume simple graph topologies (one link between two nodes) or rely on simplifying assumptions that violate ITU-T flex-grid parameters. Consequently, there is a distinct necessity for a novel, stateful approach—such as a dedicated bitmap-based parallel controller—to bridge this gap between theoretical RSA and practical, scalable deployment in modern multi-link infrastructures.